\documentclass[10pt]{report}
\usepackage{refalg}

\pagestyle{fancy}

\let\Chaptermark\chaptermark
\def\chaptermark#1{\def\Chaptername{#1}\Chaptermark{#1}}

\setlength{\parindent}{0pt}

\renewcommand{\abstractname}{Preface}

\author{ARessegetes Stery}
\title{MATH594 Algebra II}

\begin{document}

\lhead{\Chaptername}

\maketitle

\centerline{
\begin{minipage}{0.8\textwidth}
    \begin{normalsize}
        \begin{abstract}
            The notes are taken for MATH594 Algebra II by Prof. \href{https://websites.umich.edu/~mmustata/}{Mircea Mustaţă} in 2024 Winter at University of Michigan. The contents covered are mainly on basic group theory, with its application in representation theory; and field extension with Galois theory. Here I appreciate the kind help of Prof. Mustaţă throughout the whole semester.

            \vspace{1em}
            Some contents are marked with asterisk $(^*)$. These are contents excluded from testing, but out of personal interest I have recorded it. There are also some minor results that I added from the original course contents (which indicates that there may be typos or errors in the notes). Revisions, or pull requests to this repository, are more than welcomed. 

            \vspace{1em}
            For notes separated for each chapter one may consult \href{https://github.com/ARessegetesStery/MATH594-Notes}{this repository}.
        \end{abstract}
    \end{normalsize}
\end{minipage}
}
    
\tableofcontents

\setlength{\parskip}{1em}

\chapter{Group Theory}
\input{chapters/c1.tex}

\chapter{Representation of Finite Groups}
\input{chapters/c2.tex}

\chapter{Finite Fields and Galois Theory}
\input{chapters/c3.tex}

\end{document}