\documentclass[10pt]{report}
\usepackage{refalg}

\pagestyle{fancy}

\let\Chaptermark\chaptermark
\def\chaptermark#1{\def\Chaptername{#1}\Chaptermark{#1}}

\renewcommand{\abstractname}{Preface}

\author{ARessegetes Stery}
\title{MATH593 Algebra I}

\begin{document}

\lhead{\Chaptername}

\maketitle

\centerline{
\begin{minipage}{0.8\textwidth}
    \begin{normalsize}
        \begin{abstract}
            The notes are taken for MATH593 Algebra I by Prof. \href{https://websites.umich.edu/~mmustata/}{Mircea Mustaţă} in 2023 Fall at University of Michigan. The contents covered are mainly on module theory, with a bit of extension into Commutative Algebra, Category Theory and Homological Algebra. Here I appreciate the kind help of Prof. Mustaţă throughout the whole semester.

            \vspace{1em}
            Some contents are marked with asterisk $(^*)$. These are contents excluded from testing, but out of personal interest I have recorded it. There are also some minor results that I added from the original course contents (which indicates that there may be typos or errors in the notes). Revisions, or pull requests to this repository, are more than welcomed. 

            \vspace{1em}
            For notes separated for each chapter one may consult \href{https://github.com/ARessegetesStery/MATH593-Notes}{this repository}.
        \end{abstract}
    \end{normalsize}
\end{minipage}
}
    
\tableofcontents

\setlength{\parskip}{1em}

\chapter{Ring}
\input{chapters/c1.tex}

\chapter{Module}
\input{chapters/c2.tex}

\chapter{Linear Algebra on a Ring}
\input{chapters/c3.tex}

\chapter{Categories and Functors}
\input{chapters/c4.tex}

\chapter{Tensor Product}
\input{chapters/c5.tex}

\chapter{Introduction to Homological Algebra}
\input{chapters/c6.tex}

\chapter{Multilinear Algebra}
\input{chapters/c7.tex}

\end{document}